Causal Bandits: Learning Good Interventions via Causal Inference

### 2016-02-02 causal-bandit/icml-2015/introduction.tex
Tweaks

In \S\ref{sec:parallel} we analyse a specific family of causal bandit problems which we call \emph{parallel bandit} problems.
We  upper bound on the (simple) regret (Theorem~\ref{thm:}) for a general, fully observable causal model (TODO?)
\todom{Finish describing new structure}

>>>>

In \S\ref{sec:simple-regret} we analyse a specific family of causal bandit problems which we call \emph{parallel bandit} problems.
We propose a causal best arm identification algorithm (Algorithm~\ref{alg:simple}) for this problem and show it is optimal (Theorems~\ref{thm:uq-simple} \& \ref{thm:lower}).

for a general, fully observable causal model (TODO?)
\todom{Finish describing new structure}

### 2016-02-02 causal-bandit/icml-2015/setup.tex
Tweaks

The learning goal we will consider and analyse is identifying the intervention with highest expected reward by round $T$. 
This objective is most appropriate when the learning agent has a fixed learning budget after which its policy will be fixed indefintely. 
\todom{Citation}
We denote the expected reward for the intervention $do(\vec{X} = \vec{x})$ by $\mu_{\vec{x}} := \E{Y | do(\vec{X} = \vec{x})}$ and the optimal expected reward by $\mu^* := \max_{\vec{x}} \mu_{\vec{x}}$.
Performance for the first goal can be measured by \emph{simple regret}, \ie, the gap between the optimal expected reward and the intervention chosen by the learner on round $T$:
\eqn{
\label{eq:regret-simple}
\simpleregret = \mu^* - \E{\mu_{\vec{x}_T}}\,.
}

>>>>

The learning goal we will consider and analyse is identifying the intervention with highest expected reward by round $T$.
This is sometimes refered to as a ``pure exploration''~\citep{Bubeck2009} or ``best-arm identification'' problem~\citep{Gabillon2012}.
This objective is most appropriate when the learning agent has a fixed learning budget after which its policy will be fixed indefintely. 
We denote the expected reward for the intervention $a = do(\vec{X} = \vec{x})$ by $\mu_{a} := \E{Y | do(\vec{X} = \vec{x})}$ and the optimal expected reward by $\mu^* := \max_{a\in\actions} \mu_{a}$.
Performance for the first goal can be measured by \emph{simple regret}, \ie, the gap between the optimal expected reward and the intervention $a_T$ chosen by the learner on round $T$:
\begin{equation}
\label{eq:regret-simple}
	\simpleregret = \mu^* - \E{\mu_{a_T}}.
\end{equation}

### 2016-02-02 causal-bandit/icml-2015/simple-regret.tex
Tweaks

In this section we develop an algorithm for achieving the optimal simple regret in a special case of a causal bandit problem.

@@

The causal model for this class of problems has $N$ binary variables $\{ X_1, \ldots, X_N \}$ where each $X_i \in \{0,1\}$ are independent causes of a reward variable $Y \in \R$, as shown in Figure~\ref{fig:causalStructure}.
All variables are observable (\ie, $\mathcal{O} = \{Y, X_1, \ldots, X_N\}$) and the set of allowable actions are all size 0 and size 1 interventions: 
\[
	\mathcal{A} = \{do()\} \cup \{ do(X_i = j) \colon i \in \{1, \ldots, N\}, j \in \{0,1\}\}.
\]
In the farming example from the introduction, $X_1$ might represent temperature (\eg, $X_1=0$ for low and $X_1=1$ for high). 
In this case, the interventions $do(X_1 = 0)$ and $do(X_1 = 1)$ might represent the use of shades or heat lamps to keep the temperature low or high, respectively.

@@

\begin{algorithm}[h]
\caption{Causal Best Arm Identification}\label{alg:simple}
\begin{algorithmic}[1]
\STATE {\bf Input:} Total rounds $T$, causal graph $\mathcal{G}$, allowable interventions $\actions$. 
\FOR{$t \in 1,\ldots,(T - 1) / 2$}
\STATE Perform empty intervention $do()$; observe $\vec{x}_t$ and $r_t$
\ENDFOR
\FOR{$a = do(X_i = x) \in \actions$}
\STATE Count times $a$ was seen $T_a = \sum_{t=1}^{T/2} \ind{x_{t,i} = x}$
\STATE Estimate reward $\hat{\mu}_a = \frac{1}{T_a} \sum_{t=1}^{T/2} \ind{x_{t,i} = x} r_t$
\STATE Estimate probabilities $\hat{p}_a = \frac{2 T_a}{T-1}$ and $\hat{V}_a = \frac{1}{\hat{p}_a}$
\ENDFOR
\STATE Compute $\hat{m} = m(\hat{\vec{V}})$ and $A = \set{a \in \actions \colon \hat{V}_a \geq \hat{m}}$.
\STATE Let $T_A := \frac{T-1}{2 |A|}$ be times to sample each $a\in\actions$.
\FOR{$a = do(X_i = x) \in A$}
\FOR{$t \in 1,\ldots,T_A$}
\STATE Intervene with $a$ and observe $r_t$
\ENDFOR
\STATE Re-estimate $\hat{\mu}_a = \frac{1}{T_A} \sum_{t=1}^{T_A} r_t$
\ENDFOR
\RETURN estimated optimal $\hat{a}^* \in \argmax_{a\in\actions} \hat{\mu}_a$
\end{algorithmic}
\end{algorithm}

>>>>

In this section we propose and analyse an algorithm for achieving the optimal simple regret in a special case of a causal bandit problem.

@@

The causal model for this class of problems has $N$ binary variables $\{ X_1, \ldots, X_N \}$ where each $X_i \in \{0,1\}$ are independent causes of a reward variable $Y \in \R$, as shown in Figure~\ref{fig:causalStructure}.
All variables are observable and the set of allowable actions are all size 0 and size 1 interventions: 
\[
	\mathcal{A} = \{do()\} \cup \{ do(X_i = j) \colon i \in \{1, \ldots, N\}, j \in \{0,1\}\}.
\]
In the farming example from the introduction, $X_1$ might represent temperature (\eg, $X_1=0$ for low and $X_1=1$ for high). 
In this case, the interventions $do(X_1 = 0)$ and $do(X_1 = 1)$ might represent the use of shades or heat lamps to keep the temperature low or high, respectively.

@@

\begin{algorithm}[h]
\caption{Causal Best Arm Identification}\label{alg:simple}
\begin{algorithmic}[1]
\STATE {\bf Input:} Total rounds $T$, causal graph $\mathcal{G}$, allowable interventions $\actions$. 
\FOR{$t \in 1,\ldots,(T - 1) / 2$}
\STATE Perform empty intervention $do()$; observe $\vec{x}_t$ and $r_t$
\ENDFOR
\FOR{$a = do(X_i = x) \in \actions$}
\STATE Count times $X_i = x$ seen: $T_a = \sum_{t=1}^{T/2} \ind{x_{t,i} = x}$
\STATE Estimate reward: $\hat{\mu}_a = \frac{1}{T_a} \sum_{t=1}^{T/2} \ind{x_{t,i} = x} r_t$
\STATE Estimate probabilities: $\hat{p}_a = \frac{2 T_a}{T-1}$ and $\hat{V}_a = \frac{1}{\hat{p}_a}$
\ENDFOR
\STATE Compute $\hat{m} = m(\hat{\vec{V}})$ and $A = \set{a \in \actions \colon \hat{V}_a \geq \hat{m}}$.
\STATE Let $T_A := \frac{T-1}{2 |A|}$ be times to sample each $a\in A$.
\FOR{$a = do(X_i = x) \in A$}
\FOR{$t \in 1,\ldots,T_A$}
\STATE Intervene with $a$ and observe $r_t$
\ENDFOR
\STATE Re-estimate $\hat{\mu}_a = \frac{1}{T_A} \sum_{t=1}^{T_A} r_t$
\ENDFOR
\RETURN estimated optimal $\hat{a}^* \in \argmax_{a\in\actions} \hat{\mu}_a$
\end{algorithmic}
\end{algorithm}
